### 2015-12-16 ada-ftrl-proofs.tex
Improve wording

\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
We prove the lemma by induction on $T$.  The base case $T=0$ is trivial. For $T
\ge 1$, we have
$$
\Delta_T
\le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\}
\le \sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
$$
where the first inequality follows from Lemma~\ref{lemma:gap-recurrence}, and
the second inequality from the induction hypothesis and the fact that the
function $f(x) = x + \min\{a_T, a_T^2/x\}$ is increasing on $[0,\infty)$.  It
remains to prove that
$$
\sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
If $a_T = 0$, the inequality clearly holds. Otherwise, $a_T$ is positive and we
divide through by it and make substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1}
a_t^2}}{a_T}$, which leads to
$$
z\sqrt{3} + \min\left\{1,\frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
$$
which can be easily checked by considering separately the cases $z \in
[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
\end{proof}

>>>>

\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
We prove the lemma by induction on $T$.  The base case $T=0$ is trivial. For $T
\ge 1$, we have
$$
\Delta_T
\le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\}
\le \sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
$$
where the second inequality from the induction hypothesis and the fact that the
function $f(x) = x + \min\{a_T, a_T^2/x\}$ is increasing on $[0,\infty)$.  It
remains to prove that
$$
\sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
If $a_T = 0$, the inequality clearly holds. Otherwise, $a_T$ is positive and we
divide through by it and make substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1}
a_t^2}}{a_T}$, which leads to
$$
z\sqrt{3} + \min\left\{1,\frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
$$
which can be easily checked by considering separately the cases $z \in
[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
\end{proof}

### 2015-12-16 mirror-descent.tex
Mention Pistol and fixes
Improve wording

Replacing $\frac{1}{2}\|w\|_2^2$ in equation
\eqref{equation:ada-grad-regularizer} with an arbitrary convex function $R:K
\to \R$ yields
\begin{equation}
\label{equation:scale-free-mirror-descent}
R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
\end{equation}
We call the resulting algorithm \textsc{Scale-Free MD} with
regularizer $R$. Similar to \textsc{SOLO FTRL}, the regularizer
\eqref{equation:scale-free-mirror-descent} does not uniquely define
the \textsc{MD} minimizer
$w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
w_t) \right)$ when $\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}$ is zero.
This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
zero. In this case, we define
$w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
$w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.
Similarly, $w_1 = \argmin_{w \in K} R(w)$.

The theorem below upper bounds the regret of \textsc{Scale-Free MD} (see also
\cite{Duchi-Hazan-Singer-2011,Duchi-Shalev-Shwartz-Singer-Tewari-2010,Rakhlin-Sridharan-2013}).
Proof is in~\ref{section:mirror-descent-proofs}.

@@

\subsection{Lower Bounds for Mirror Descent}
\label{subsection:mirror-descent-lower-bound}

@@

The first counter-example is stated as
Theorem~\ref{theorem:first-counter-example} below. The decision set is the
whole space $K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note
that $R(w)$ is $1$-strongly convex with respect to $\|\cdot\|_2$ and the dual
norm of $\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
the first vector of the standard unit basis.  On such sequence, Zinkevich's
\textsc{GIGA} and both versions of \textsc{AdaGrad MD} have predictions
identical to \textsc{Scale-Free MD}. Hence the lower bound applies to all four
algorithms.

@@

The second counter-example is stated as
Theorem~\ref{theorem:second-counter-example} below.  The decision set is the
$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
\sum_{j=1}^d w_j = 1 \}$ and the regularizer is the negative entropy $R(w) =
\sum_{j=1}^d w_j \ln w_j$.  Negative entropy is $1$-strongly convex with
respect to $\|\cdot\|_1$ and the dual norm of $\|\cdot\|_1$ is
\mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman divergence is the
Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{j=1}^d u_j \ln(u_j/v_j)$.
Note that despite that negative entropy is upper- and lower-bounded,
Kullback-Leibler divergence can be arbitrarily large.

>>>>

Replacing $\frac{1}{2}\|w\|_2^2$ in equation
\eqref{equation:ada-grad-regularizer} with an arbitrary strongly convex
function $R:K \to \R$ yields
\begin{equation}
\label{equation:scale-free-mirror-descent}
R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
\end{equation}
We call the resulting algorithm \textsc{Scale-Free MD} with
regularizer $R$. Similar to \textsc{SOLO FTRL}, the regularizer
\eqref{equation:scale-free-mirror-descent} does not uniquely define
the \textsc{MD} minimizer
$w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
w_t) \right)$ when $\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}$ is zero.
This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
zero. In this case, we define
$w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
$w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.
Similarly, $w_1 = \argmin_{w \in K} R(w)$.

The theorem below upper bounds the regret of \textsc{Scale-Free MD} (see also
\cite{Duchi-Hazan-Singer-2011, Duchi-Shalev-Shwartz-Singer-Tewari-2010,
Rakhlin-Sridharan-2013}).  Proof is in~\ref{section:mirror-descent-proofs}.

@@

\subsection{Lower Bounds for Scale-Free Mirror Descent}
\label{subsection:mirror-descent-lower-bound}

@@

The first counter-example is stated as
Theorem~\ref{theorem:first-counter-example} below; proof is
in~\ref{section:mirror-descent-proofs}. The decision set is the whole space
$K=V$ and the regularizer is $R(w) = \frac{1}{2}\|w\|_2^2$. Note that $R(w)$ is
$1$-strongly convex with respect to $\|\cdot\|_2$ and the dual norm of
$\|\cdot\|_2$ is $\|\cdot\|_2$. The corresponding Bregman divergence is
$\Breg_R(u,v) = \frac{1}{2}\|u-v\|_2^2$. The counter-example constructs
sequence of unit-norm loss vectors in the one-dimensional subspace spanned by
the first vector of the standard unit basis.  On such sequence, Zinkevich's
\textsc{GIGA} and both versions of \textsc{AdaGrad MD} have predictions
identical to \textsc{Scale-Free MD}. Hence the lower bound applies to all four
algorithms.

@@

The second counter-example is stated as
Theorem~\ref{theorem:second-counter-example} below; proof is
in~\ref{section:mirror-descent-proofs}.  The decision set is the
$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
\sum_{j=1}^d w_j = 1 \}$ and the regularizer is the negative entropy $R(w) =
\sum_{j=1}^d w_j \ln w_j$.  Negative entropy is $1$-strongly convex with
respect to $\|\cdot\|_1$ and the dual norm of $\|\cdot\|_1$ is
\mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman divergence is the
Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{j=1}^d u_j \ln(u_j/v_j)$.
Note that despite that negative entropy is upper- and lower-bounded,
Kullback-Leibler divergence can be arbitrarily large.
